---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# **Coral reefs benthic monitoring: spatio-temporal distribution, methods and stakeholders**


**This repository contains code associated to the article:**

Wicquart, J., Prasil-Delaval, N., Obura, D., Gudka, M., Logan, M., Souter, D., Staub, F., Towle, E.K., Burt, J.A., Padovani Ferreira, B., Kimura, T., Chan, S., Tun, K., Ibrahim, N., Amer, M.,  Reyes Bonilla, H., Smith, F.,  GCRMN Consortium, Planes, S. (2022). Coral reefs benthic monitoring: spatio-temporal distribution, methods and stakeholders. _in prep_.

## 1. Abstract



## 2. How to download this project?

On the project main page on GitHub, click on the green button `Code` and then click on `Download ZIP`

## 3. Context

*[To be completed]*

## 4. Description of the project

### 4.1 Project organization

* :open_file_folder: `R`
  + :open_file_folder: `functions` Functions used in the two main scripts
    - `graphical_par.R` Graphical parameters
    - `theme_graph.R` Theme for graphs
    - `theme_map.R` Theme for maps (primary)
    - `theme_map2.R` Theme for maps (secondary)
  + `01_reefs-ecoregion.R` Filter ecoregions containing coral reefs
  + `02_analysis.Rmd` Main analysis script

### 4.2 Data description

Four datasets were used for the analyses:

* Base map datasets downloaded on [Natural Earth Data](https://www.naturalearthdata.com/downloads/10m-physical-vectors/).

* Marine Ecoregions of the World (MEOW) dataset from [Spalding *et al*, 2007](https://academic.oup.com/bioscience/article/57/7/573/238419).

* Distribution of coral reefs of the world from [UNEP-WCMC](https://data.unep-wcmc.org/datasets/1).

* Synthetic dataset used for the [*Status of Coral Reefs of the World: 2020*](https://gcrmn.net/2020-report/) report.

## 5. How to report issues?

Please report any bugs or issues [HERE](https://github.com/JWicquart/monitoring_reefs/issues).


## 6. Reproducibility parameters

```{r echo=FALSE, message=FALSE}

# 1. Print session info ----

devtools::session_info(c("sf", "tidyverse", "magrittr", "readxl", "ggstream",
                         "patchwork", "ggtext", "glue", "ggrepel", "FactoMineR", "RColorBrewer", "ggforce"))

```
