---
title: "Analyses"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: "cosmo"
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Load packages and data

```{r}

# 1. Load packages ----

library(magrittr) # To use special pipes
library(tidyverse) # Core tidyverse packages
library(sf)
sf_use_s2(FALSE) # Switch from S2 to GEOS
library(ggstream)
library(cowplot)
library(ggsflabel)
library(ggtext)
library(glue)
library(ggrepel)
library(FactoMineR) # For PCA
library(RColorBrewer)
library(ggforce)# For geom_circle

# 2. Source functions ----

source("functions/graphical_par.R")
source("functions/theme_map.R")
source("functions/theme_map2.R")
source("functions/theme_graph.R")
theme_set(theme_graph())

# 3. Load MEOW data ----

load(file = "../data/04_meow-with-coral-reefs.RData")

# 4. Load 2020 GCRMN data ----

data_gcrmn <- read.csv2(file = "../data/03-merge_all_all_all_benthos_NA.csv") %>% 
  filter(!(DatasetID %in% c("XLCA1", "XLCA2", "XLCA3", "XLCA4", "XLCA5"))) %>% 
  drop_na(Latitude, Longitude) %>% 
  # Add data sources
  left_join(., read.csv2("../data/05_data-sources.csv"))

# 5. Remove sites falling outside MEOW containing coral reefs ----

data_gcrmn <- data_gcrmn %>% 
  mutate(lat = Latitude, 
         long = Longitude) %>% 
  st_as_sf(., coords = c("long", "lat"), crs = 4326) %>% 
  st_join(., data_meowreefs) %>% 
  drop_na(ECOREGION) %>% 
  st_drop_geometry() %>% 
  select(-ECOREGION:-SHAPE_Area)

# 6. Summarize by duration ----

data_gcrmn_sites <- data_gcrmn %>% 
  group_by(Area, Country, Location, Site, Latitude, Longitude) %>% 
  drop_na(Year) %>% 
  summarise(interval_years = max(Year, na.rm = TRUE) - min(Year, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(interval_class = cut(interval_years, 
                              breaks = c(-Inf, 1, 5, 10, 15, Inf),
                              labels = c("1", "2-5", "6-10", "11-15", ">15"))) %>% 
  st_as_sf(., coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform(., crs = crs_selected)

```

# 2. Maps

## 2.1 Load data

```{r}

# 1. Load background maps ----

data_map <- read_sf("../data/00_natural-earth-data/ne_10m_land/ne_10m_land.shp") %>% 
  st_transform(crs = 4326) %>% 
  st_transform(crs = crs_selected)

# 2. Create the border of background map ----

lats <- c(90:-90, -90:90, 90)
longs <- c(rep(c(180, -180), each = 181), 180)

background_map_border <- list(cbind(longs, lats)) %>%
  st_polygon() %>%
  st_sfc(crs = 4326) %>% 
  st_sf() %>%
  st_transform(crs = crs_selected)

```

## 2.2 Global map MEOW

```{r}

# 1. Make the map ----

ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_meowreefs, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  guides(colour = guide_legend(override.aes = list(size = 5))) +
  scale_color_manual(values = palette_duration) +
  theme_map() +
  theme(legend.title = element_blank())

# 2. Save the map ----

ggsave("../figs/01_global-meow-with-coral-reefs.png", height = 6, width = 10)

```

## 2.3 Global map duration

```{r}

# 1. Make the map ----

ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_meowreefs, fill = "#ecf0f1") +
  # Site of benthic cover data
  geom_sf(data = data_gcrmn_sites, aes(color = interval_class)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  guides(colour = guide_legend(title.position = "top", title.hjust = 0.5, override.aes = list(size = 5))) +
  scale_color_manual(values = palette_duration) +
  theme_map() +
  labs(col = "Monitoring duration (years)")

# 2. Save the map ----

ggsave("../figs/01_global-map-site-duration.png", height = 6, width = 10)

```

## 2.4 Regional maps duration

```{r}

# 1. Transform CRS for data map ----

data_map <- data_map %>% 
  st_transform(., crs = 4326)

# 2. Transform CRS for data GCRMN sites ----

data_gcrmn_sites <- data_gcrmn_sites %>% 
  st_transform(., crs = 4326)

```

### Australia

```{r}

# 1. Create the buffer ----

data_buffer <- st_sf(coord = st_sfc(st_point(c(133, -25))), crs = 4326) %>%
  st_buffer(dist = 28)

# 2. Intersection based on buffer for data_map ----

data_map_region <- st_intersection(data_buffer, data_map)

# 3. Select MEOW for the region ----

data_meowreefs_region <- data_meowreefs %>% 
  filter(ECO_CODE_X %in% c(141, 140, 142, 150, 143, 202, 203, 204, 205, 206, 207, 208, 209, 211, 210, 145, 144))

# 4. Filter sites ----

data_sites_region <- st_intersection(data_meowreefs_region, data_gcrmn_sites)

# 5. Make the map ----

ggplot() +
  # First data buffer (for ocean color background)
  geom_sf(data = data_buffer, fill = col_background_map) +
  # MEOW
  geom_sf(data = data_meowreefs_region, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map_region, fill = col_fill_map, col = col_color_map) +
  # GCRMN sites
  geom_sf(data = data_sites_region, aes(color = interval_class), show.legend = FALSE) +
  #geom_sf_text_repel(data = data_meowreefs_region, aes(label = ECO_CODE_X), size = 2, family = font_choose_map) +
  scale_color_manual(values = palette_duration) +
  # Second data buffer (for borders)
  geom_sf(data = data_buffer, fill = NA, size = 0.75) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "Australia")
 
# 6. Save the map ----

ggsave("../figs/06_meow-australia.png", dpi = 600, width = 4, height = 5)

```

### Caribbean

```{r}

# 1. Create the buffer ----

data_buffer <- st_sf(coord = st_sfc(st_point(c(-85, 18))), crs = 4326) %>%
  st_buffer(dist = 35)

# 2. Intersection based on buffer for data_map ----

data_map_region <- st_intersection(data_buffer, data_map)

# 3. Select MEOW for the region ----

data_meowreefs_region <- data_meowreefs %>% 
  filter(ECO_CODE_X %in% c(69, 43, 70, 63, 65, 64, 66, 67, 62, 59, 60, 61, 164, 165, 166, 167,
                           168, 169, 170, 171, 172, 173, 174, 175, 68, 42))

# 4. Filter sites ----

data_sites_region <- st_intersection(data_meowreefs_region, data_gcrmn_sites)

# 5. Make the map ----

ggplot() +
  # First data buffer (for ocean color background)
  geom_sf(data = data_buffer, fill = col_background_map) +
  # MEOW
  geom_sf(data = data_meowreefs_region, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map_region, fill = col_fill_map, col = col_color_map) +
  # GCRMN sites
  geom_sf(data = data_sites_region, aes(color = interval_class), show.legend = FALSE) +
  #geom_sf_text_repel(data = data_meowreefs_region, aes(label = ECO_CODE_X), size = 2, family = font_choose_map) +
  scale_color_manual(values = palette_duration) +
  # Second data buffer (for borders)
  geom_sf(data = data_buffer, fill = NA, size = 0.75) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "Caribbean and Eastern Pacific")
 
# 6. Save the map ----

ggsave("../figs/06_meow-caribbean-and-eastern-pacific.png", dpi = 600, width = 4, height = 5)

```

### Red Sea and Gulf

```{r}

# 1. Create the buffer ----

data_buffer <- st_sf(coord = st_sfc(st_point(c(47, 20))), crs = 4326) %>%
  st_buffer(dist = 20)

# 2. Intersection based on buffer for data_map ----

data_map_region <- st_intersection(data_buffer, data_map)

# 3. Select MEOW for the region ----

data_meowreefs_region <- data_meowreefs %>% 
  filter(ECO_CODE_X %in% c(87, 88, 89, 90, 91, 92))

# 4. Filter sites ----

data_sites_region <- st_intersection(data_meowreefs_region, data_gcrmn_sites)

# 5. Make the map ----

ggplot() +
  # First data buffer (for ocean color background)
  geom_sf(data = data_buffer, fill = col_background_map) +
  # MEOW
  geom_sf(data = data_meowreefs_region, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map_region, fill = col_fill_map, col = col_color_map) +
  # GCRMN sites
  geom_sf(data = data_sites_region, aes(color = interval_class), show.legend = FALSE) +
  #geom_sf_text_repel(data = data_meowreefs_region, aes(label = ECO_CODE_X), size = 2, family = font_choose_map) +
  scale_color_manual(values = palette_duration) +
  # Second data buffer (for borders)
  geom_sf(data = data_buffer, fill = NA, size = 0.75) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "Red Sea and Gulf")
 
# 6. Save the map ----

ggsave("../figs/06_meow-red-sea-and-gulf.png", dpi = 600, width = 4, height = 5)

```

### Western Indian Ocean

```{r}

# 1. Create the buffer ----

data_buffer <- st_sf(coord = st_sfc(st_point(c(50, -12))), crs = 4326) %>%
  st_buffer(dist = 27)

# 2. Intersection based on buffer for data_map ----

data_map_region <- st_intersection(data_buffer, data_map)

# 3. Select MEOW for the region ----

data_meowreefs_region <- data_meowreefs %>% 
  filter(ECO_CODE_X %in% c(93, 94, 95, 96, 97, 98, 99, 100, 101, 102, 193))

# 4. Filter sites ----

data_sites_region <- st_intersection(data_meowreefs_region, data_gcrmn_sites)

# 5. Make the map ----

ggplot() +
  # First data buffer (for ocean color background)
  geom_sf(data = data_buffer, fill = col_background_map) +
  # MEOW
  geom_sf(data = data_meowreefs_region, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map_region, fill = col_fill_map, col = col_color_map) +
  # GCRMN sites
  geom_sf(data = data_sites_region, aes(color = interval_class), show.legend = FALSE) +
  #geom_sf_text_repel(data = data_meowreefs_region, aes(label = ECO_CODE_X), size = 2, family = font_choose_map) +
  scale_color_manual(values = palette_duration) +
  # Second data buffer (for borders)
  geom_sf(data = data_buffer, fill = NA, size = 0.75) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "Western Indian Ocean")
 
# 6. Save the map ----

ggsave("../figs/06_meow-western-indian-ocean.png", dpi = 600, width = 4, height = 5)

```

### Brazil

```{r}

# 1. Create the buffer ----

data_buffer <- st_sf(coord = st_sfc(st_point(c(-35, -12))), crs = 4326) %>%
  st_buffer(dist = 18)

# 2. Intersection based on buffer for data_map ----

data_map_region <- st_intersection(data_buffer, data_map)

# 3. Select MEOW for the region ----

data_meowreefs_region <- data_meowreefs %>% 
  filter(ECO_CODE_X %in% c(73, 74, 75, 76, 77, 180))

# 4. Filter sites ----

data_sites_region <- st_intersection(data_meowreefs_region, data_gcrmn_sites)

# 5. Make the map ----

ggplot() +
  # First data buffer (for ocean color background)
  geom_sf(data = data_buffer, fill = col_background_map) +
  # MEOW
  geom_sf(data = data_meowreefs_region, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map_region, fill = col_fill_map, col = col_color_map) +
  # GCRMN sites
  geom_sf(data = data_sites_region, aes(color = interval_class), show.legend = FALSE) +
  #geom_sf_text_repel(data = data_meowreefs_region, aes(label = ECO_CODE_X), size = 2, family = font_choose_map) +
  scale_color_manual(values = palette_duration) +
  # Second data buffer (for borders)
  geom_sf(data = data_buffer, fill = NA, size = 0.75) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "Brazil")
 
# 6. Save the map ----

ggsave("../figs/06_meow-brazil.png", dpi = 600, width = 4, height = 5)

```

### South Asia

```{r}

# 1. Create the buffer ----

data_buffer <- st_sf(coord = st_sfc(st_point(c(77, 10))), crs = 4326) %>%
  st_buffer(dist = 25)

# 2. Intersection based on buffer for data_map ----

data_map_region <- st_intersection(data_buffer, data_map)

# 3. Select MEOW for the region ----

data_meowreefs_region <- data_meowreefs %>% 
  filter(ECO_CODE_X %in% c(103, 104, 105, 106, 107))

# 4. Filter sites ----

data_sites_region <- st_intersection(data_meowreefs_region, data_gcrmn_sites)

# 5. Make the map ----

ggplot() +
  # First data buffer (for ocean color background)
  geom_sf(data = data_buffer, fill = col_background_map) +
  # MEOW
  geom_sf(data = data_meowreefs_region, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map_region, fill = col_fill_map, col = col_color_map) +
  # GCRMN sites
  geom_sf(data = data_sites_region, aes(color = interval_class), show.legend = FALSE) +
  #geom_sf_text_repel(data = data_meowreefs_region, aes(label = ECO_CODE_X), size = 2, family = font_choose_map) +
  scale_color_manual(values = palette_duration) +
  # Second data buffer (for borders)
  geom_sf(data = data_buffer, fill = NA, size = 0.75) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "South Asia")
 
# 6. Save the map ----

ggsave("../figs/06_meow-south-asia.png", dpi = 600, width = 4, height = 5)

```

### East Asia

```{r}

# 1. Create the buffer ----

data_buffer <- st_sf(coord = st_sfc(st_point(c(125, 4))), crs = 4326) %>%
  st_buffer(dist = 40)

# 2. Intersection based on buffer for data_map ----

data_map_region <- st_intersection(data_buffer, data_map)

# 3. Select MEOW for the region ----

data_meowreefs_region <- data_meowreefs %>% 
  filter(ECO_CODE_X %in% c(109, 110, 111, 118, 115, 116, 114, 112, 113, 117, 119, 120, 127, 133, 126,
                           128, 129, 130, 131, 139, 138, 134, 136, 137, 132, 125, 121, 47, 48, 49, 50, 51, 52))

# 4. Filter sites ----

data_sites_region <- st_intersection(data_meowreefs_region, data_gcrmn_sites)

# 5. Make the map ----

ggplot() +
  # First data buffer (for ocean color background)
  geom_sf(data = data_buffer, fill = col_background_map) +
  # MEOW
  geom_sf(data = data_meowreefs_region, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map_region, fill = col_fill_map, col = col_color_map) +
  # GCRMN sites
  geom_sf(data = data_sites_region, aes(color = interval_class), show.legend = FALSE) +
  #geom_sf_text_repel(data = data_meowreefs_region, aes(label = ECO_CODE_X), size = 2, family = font_choose_map) +
  scale_color_manual(values = palette_duration) +
  # Second data buffer (for borders)
  geom_sf(data = data_buffer, fill = NA, size = 0.75) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "East Asia")
 
# 6. Save the map ----

ggsave("../figs/06_meow-east-asia.png", dpi = 600, width = 4, height = 5)

```

### Pacific

```{r}

# 1. Create the buffer ----

data_buffer <- st_sf(coord = st_sfc(st_point(c(171.5, 10))), crs = 4326) %>%
  st_buffer(dist = 42)

# 2. Intersection based on buffer for data_map ----

data_map_region <- st_intersection(data_buffer, data_map)

data_map_region <- st_shift_longitude(data_map_region)

# 3. Select MEOW for the region ----

data_meowreefs_region <- data_meowreefs %>% 
  filter(ECO_CODE_X %in% c(124, 135, 149, 148, 122, 123, 153, 154, 147, 157, 146, 156, 155, 152))

data_meowreefs_region <- st_shift_longitude(data_meowreefs_region)

data_meowreefs_region <- data_meowreefs_region %<>% # Special pipe from magrittr
  st_buffer(0.0015) # To join polygon (remove vertical line)

# 4. Filter sites ----

data_sites_region <- st_intersection(data_meowreefs_region, st_shift_longitude(data_gcrmn_sites))

data_sites_region <- st_shift_longitude(data_sites_region)

# 5. Make the map ----

ggplot() +
  # First data buffer (for ocean color background)
  geom_sf(data = data_buffer, fill = col_background_map) +
  # MEOW
  geom_sf(data = data_meowreefs_region, fill = "#ecf0f1") +
  # Background map
  geom_sf(data = data_map_region, fill = col_fill_map, col = col_color_map) +
  # GCRMN sites
  geom_sf(data = data_sites_region, aes(color = interval_class), show.legend = FALSE) +
  #geom_sf_text_repel(data = data_meowreefs_region, aes(label = ECO_CODE_X), size = 2, family = font_choose_map) +
  scale_color_manual(values = palette_duration) +
  # Second data buffer (for borders)
  geom_sf(data = data_buffer, fill = NA, size = 0.75) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 16)) +
  labs(title = "Pacific")
 
# 6. Save the map ----

ggsave("../figs/06_meow-pacific.png", dpi = 600, width = 4, height = 5)

```

# 3. Methods used

```{r}

# 1. Filter and transform the data ----

data_method <- data_gcrmn %>% 
  filter(Year > 1980 & Year <= 2018) %>%   
  select(Latitude, Longitude, Year, Method) %>% 
  distinct(.) %>% 
  mutate(Method = str_split_fixed(Method, ",", 6)[,1],
         Method = str_to_lower(Method),
         Method = str_trim(Method, side = "both"),
         Method = str_squish(Method),
         Method = case_when(Method %in% c("line intercept transect", 
                                          "line intersect transect",
                                          "lit") ~ "LIT",
                            Method %in% c("point intercept method",
                                          "point intercept transect",
                                          "point intersect transect",
                                          "pit") ~ "PIT",
                            Method %in% c("photo-quadrat",
                                          "Photo-quadrat",
                                          "underwater photo transect") ~ "Photo-quadrat",
                            Method %in% c("visual census",
                                          "visual quadrat estimation",
                                          "vis est.") ~ "Visual census",
                            Method == "video-transect" ~ "Video transect",
                            TRUE ~ "Unknown")) %>% 
  mutate(Method = fct_relevel(Method, c("Unknown", "Photo-quadrat", "PIT", "LIT", "Video transect", "Visual census"))) %>% 
  group_by(Method, Year) %>% 
  count(name = "Abs") %>% 
  ungroup() %>% 
  complete(Method, fill = list(Abs = 0)) %>% 
  group_by(Year) %>% 
  mutate(Rel = (Abs/sum(Abs))*100)

# 2. Define label positions ----

data_method_label <- tibble(Method = c("Unknown", "Photo-quadrat", "PIT", "LIT", "Video transect", "Visual census"),
                            x = c(1995, 2012, 2003, 1989, 2001, 1990),
                            y = c(0.94, 0.75, 0.5, 0.65, 0.22, 0.22))

# 3. Make the plot ----

ggplot(data = data_method, aes(x = Year, y = Rel, fill = Method, label = Method)) +
  geom_stream(type = "proportional", show.legend = FALSE, alpha = 0.9) +
  geom_text(data = data_method_label, aes(x = x, y = y, label = Method), family = font_choose_graph) +
  # Alternative using geom_stream_label but less precise
  #geom_stream_label(type = "proportional", family = font_choose_graph) +
  #scale_fill_manual(values = c("#dadfe1", "#FFC740", "#C20008", "#FF020D", "#13AFEF", "#2abb9b")) +
  scale_fill_brewer(palette = "RdYlBu") +
  lims(x = c(1980, 2020)) +
  labs(y = "Percentage of surveys", x = "Year") +
  scale_y_continuous(labels = scales::percent)

# 4. Save the plot ----

ggsave(filename = "../figs/02_evolution-methods-used_streamplot.png", height = 4, width = 8)

# 5. Remove useless objects ----

rm(data_method, data_method_label)

```

# 4. Data sources

```{r fig.width=10, fig.height=4}

# 1. Filter and transform the data ----

data_source <- data_gcrmn %>% 
  filter(Year > 1980 & Year <= 2018) %>%   
  select(Latitude, Longitude, Year, Source) %>% 
  distinct(.) %>%
  group_by(Source, Year) %>% 
  count(name = "Abs") %>% 
  ungroup() %>% 
  complete(Source, fill = list(Abs = 0)) %>% 
  group_by(Year) %>% 
  mutate(Rel = (Abs/sum(Abs))*100)

# 2. Define label positions ----

#data_source_label <- tibble(Source = c("Unknown", "Photo-quadrat", "PIT", "LIT", "Video transect", "Visual census"),
#                            x = c(1995, 2012, 2003, 1989, 2001, 1990),
#                            y = c(0.94, 0.75, 0.5, 0.65, 0.22, 0.22))

# 3. Make the plot ----

ggplot(data = data_source, aes(x = Year, y = Rel, fill = Source, label = Source)) +
  geom_stream(type = "proportional", show.legend = TRUE, alpha = 0.9) +
  #geom_text(data = data_method_label, aes(x = x, y = y, label = Method), family = font_choose_graph) +
  # Alternative using geom_stream_label but less precise
  #geom_stream_label(type = "proportional", family = font_choose_graph) +
  #scale_fill_manual(values = c("#dadfe1", "#FFC740", "#C20008", "#FF020D", "#13AFEF", "#2abb9b")) +
  scale_fill_brewer(palette = "RdYlBu") +
  lims(x = c(1980, 2020)) +
  labs(y = "Percentage of surveys", x = "Year") +
  scale_y_continuous(labels = scales::percent)

# 4. Save the plot ----

ggsave(filename = "../figs/02_evolution-data-sources_streamplot.png", height = 4, width = 8)

```

# 5. Histograms

## 5.1 Monitoring duration

```{r fig.width=6, fig.height=5}

# 1. Transform the data ----

data_duration <- data_gcrmn_sites %>% 
  st_drop_geometry() %>% 
  group_by(interval_class) %>% 
  summarise(n_abs = n()) %>% 
  ungroup() %>% 
  mutate(n_rel = (n_abs/sum(n_abs))*100)

# 2. Make the plot ----

plot_duration <- ggplot(data = data_duration, aes(x = interval_class, y = n_rel)) +
  geom_bar(stat = "identity", aes(fill = interval_class), show.legend = FALSE, col = "black") +
  scale_fill_manual(values = palette_duration) +
  lims(y = c(0, 100)) +
  labs(x = "Monitoring duration (years)", y = "Sites (%)")

```

## 5.2 Surveys by year

```{r}

# 1. Modify the data ----

data_year <- data_gcrmn %>% 
  select(Latitude, Longitude, Year) %>% 
  distinct(.)

# 2. Make the plot ----

plot_year <- ggplot(data_year, aes(x = Year)) +
  # By default its density, width*density*100 gives percentage
  geom_histogram(binwidth = 1, aes(y = stat(width*density*100)),
                 color = col_color_graph, fill = col_fill_graph) +
  lims(y = c(0, 10)) +
  labs(x = "Year", y = "Surveys (%)")

```

## 5.3 Surveys by depth

```{r}

# 1. Modify the data ----

data_depth <- data_gcrmn %>% 
  select(Latitude, Longitude, Year, Depth) %>% 
  distinct(.)

# 2. Make the plot ----

plot_depth <- ggplot(data = data_depth, aes(x = Depth)) +
  # By default its density, width*density*100 gives percentage
  geom_histogram(binwidth = 1, aes(y = stat(width*density*100)),
                 color = col_color_graph, fill = col_fill_graph) +
  lims(x = c(0, 40), y = c(0, 30)) +
  labs(x = "Depth (m)", y = "Surveys (%)")

```

## 5.4 Plot assemblage

```{r}

# 1. Make the plot ----

plot_grid(plot_year, plot_depth, plot_duration, nrow = 1, labels = c("A", "B", "C"))

# 2. Save the plot ----

ggsave(filename = "../figs/03_combined-plots-distri-year-depth-duration.png", height = 3, width = 10)

# 3. Remove useless objects ----

rm(data_depth, plot_depth, data_year, plot_year, data_duration, plot_duration)

```

# 6. Taxonomic resolution

```{r}

# 1. Filter and transform the data ----

data_taxo <- data_gcrmn %>% 
  filter(Year > 1980 & Year <= 2018) %>%   
  select(Year, Category, Group, Family, Genus, Species) %>% 
  mutate(Tax_lvl = case_when(!(is.na(Species)) ~ "Species",
                             !(is.na(Genus)) ~ "Genus",
                             !(is.na(Family)) ~ "Family",
                             !(is.na(Group)) ~ "Group",
                             !(is.na(Category)) ~ "Category")) %>%
  mutate(Tax_lvl = fct_relevel(Tax_lvl, c("Category", "Group", "Family", "Genus", "Species"))) %>% 
  group_by(Year, Tax_lvl) %>% 
  count(name = "Abs") %>% 
  ungroup() %>% 
  complete(Tax_lvl, fill = list(Abs = 0)) %>% 
  group_by(Year) %>% 
  mutate(Rel = (Abs/sum(Abs))*100)

# 2. Define label positions ----

data_taxo_label <- tibble(Tax_lvl = c("Category", "Group", "Family", "Genus", "Species"),
                          x = c(1990, 2010, 1998, 2000, 2010),
                          y = c(0.875, 0.65, 0.4, 0.2, 0.1))

# 3. Make the plot ----

ggplot(data = data_taxo, aes(x = Year, y = Rel, fill = Tax_lvl, label = Tax_lvl)) +
  geom_stream(type = "proportional", show.legend = FALSE) +
  geom_text(data = data_taxo_label, aes(x = x, y = y, label = Tax_lvl), family = font_choose_graph) +
  # Alternative using geom_stream_label but less precise
  #geom_stream_label(type = "proportional", family = font_choose_graph) +
  scale_fill_manual(values = c("#FFC740", "#C20008", "#FF020D", "#13AFEF", "#2abb9b")) +
  lims(x = c(1980, 2020)) +
  labs(y = "Percentage of surveys", x = "Year")

# 4. Save the plot ----

ggsave(filename = "../figs/04_evolution-taxo-resolution_streamplot.png", height = 4, width = 8)

# 5. Remove useless objects ----

rm(data_taxo, data_taxo_label)

```

# 7. Comparison between ecoregions

## 7.1 Calculation of descriptors

```{r}

# 1. Load and transform Marine Ecoregions of the World (MEOW) data ----

data_meow <- read_sf("../data/02_marine-ecoregions-of-the-world/Marine_Ecoregions_Of_the_World__MEOW_.shp") %>% 
  st_transform(crs = 4326) %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES")) %>% 
  st_make_valid()

# 2. Join MEOW and GCRMN data ----

data_gcrmn_meow <- data_gcrmn %>% 
  drop_na(Latitude, Longitude) %>% 
  st_as_sf(., coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_intersection(., data_meow) %>% 
  mutate(geometry2 = str_remove_all(geometry, "c|\\(|\\)"),
         Longitude = as.numeric(str_split_fixed(geometry2, ",", 2)[,1]),
         Latitude = as.numeric(str_split_fixed(geometry2, ",", 2)[,2])) %>% 
  st_drop_geometry() %>% 
  select(-geometry2)

# 3. Calculation of sites ----

data_descriptors <- data_gcrmn_meow %>% 
  select(Latitude, Longitude, ECOREGION) %>% 
  distinct() %>% 
  group_by(ECOREGION) %>% 
  summarise(n_sites = n())

# 4. Calculation of observations ----

data_descriptors <- data_gcrmn_meow %>% 
  group_by(ECOREGION) %>% 
  summarise(n_obs = n()) %>% 
  left_join(data_descriptors, .)
  
# 5. Calculation of surveys ----

data_descriptors <- data_gcrmn_meow %>% 
  select(Latitude, Longitude, Year, ECOREGION) %>% 
  distinct() %>% 
  group_by(ECOREGION) %>% 
  summarise(n_surveys = n()) %>% 
  left_join(data_descriptors, .)

# 6. Calculation of monitoring duration ----

data_descriptors <- data_gcrmn_meow %>% 
  group_by(Area, Country, Location, Site, Latitude, Longitude, ECOREGION) %>% 
  summarise(interval_years = max(Year, na.rm = TRUE) - min(Year, na.rm = TRUE)) %>%
  drop_na(interval_years) %>% 
  ungroup() %>% 
  mutate(interval_class = cut(interval_years, 
                              breaks = c(-Inf, 1, 5, 10, 15, Inf),
                              labels = c("n_1", "n_2_5", "n_6_10", "n_11_15", "n_15"))) %>% 
  group_by(ECOREGION, interval_class) %>%
  summarise(n = n()) %>% 
  pivot_wider(names_from = interval_class, values_from = n, values_fill = 0) %>% 
  left_join(data_descriptors, .)

# 7. Number of years with data ---- 

data_descriptors <- data_gcrmn_meow %>% 
  select(ECOREGION, Year) %>% 
  distinct() %>% 
  group_by(ECOREGION) %>% 
  summarise(n_years = n(),
            first_year = min(Year)) %>% 
  ungroup() %>% 
  left_join(data_descriptors, .)

# 8. Add reef surfaces ----

load("../data/04_reef-surface-by-ecoregion.RData")

data_descriptors <- left_join(data_surface, data_descriptors) %>% 
  # Replace NA with 0 for some variables
  mutate_at(vars(starts_with("n_")), ~replace_na(., 0))

# 9. Calculate relative frequency ----

data_descriptors <- data_descriptors %>% 
  pivot_longer("area":"n_years", names_to = "var_names", values_to = "abs") %>% # Select variables to shift in relative freq here
  group_by(var_names) %>% 
  mutate(total = sum(abs)) %>% 
  ungroup() %>% 
  group_by(ECOREGION, var_names) %>% 
  mutate(rel = (abs/total)*100) %>% 
  ungroup() %>% 
  select(-total) %>% 
  pivot_wider(names_from = "var_names", values_from = c("abs", "rel"), names_glue = "{var_names}_{.value}")

```

```{r}

# 1. Import data on coral reefs distribution ----

data_reefs <- read_sf("./../data/01_reefs-distribution-wri/reef_500_poly.shp") %>% 
  st_transform(crs = 4326) %>% 
  st_wrap_dateline(options = c("WRAPDATELINE=YES")) %>% 
  st_make_valid()

# 2. Create a grid cells of 0.5° ----

grid_reefs <- st_make_grid(data_reefs, cellsize = 1) %>% 
  st_as_sf() %>% 
  st_filter(., data_reefs) # Filter grid cells containing coral reefs

# 3. Make intersection with MEOW ----

grid_reefs <- st_intersection(grid_reefs, data_meow)

# 4. Count number of polygon by MEOW ---- 

pixel_total <- grid_reefs %>% 
  st_drop_geometry() %>% 
  group_by(ECOREGION, ECO_CODE_X) %>% 
  count(name = "n_total")

# 5. Count number of polygon with monitoring data by MEOW ---- 

pixel_monitoring <- data_gcrmn_sites %>% 
  st_transform(crs = 4326) %>% 
  st_filter(grid_reefs, .) %>% 
  st_drop_geometry() %>% 
  group_by(ECOREGION, ECO_CODE_X) %>% 
  count(name = "n_monitoring")

# 6. Join with other data descriptors ----

data_descriptors <- left_join(pixel_total, pixel_monitoring) %>% 
  mutate(n_monitoring = replace_na(n_monitoring, replace = 0),
         n_percent = (n_monitoring/n_total)*100) %>% 
  left_join(data_descriptors, .)

# 7. Transform as an sf object ----

data_descriptors <- left_join(as_tibble(data_meowreefs), data_descriptors) %>% 
  st_as_sf() %>% 
  select(-"FID":-"SHAPE_Area", -"area")

# 8. Map of grid reefs to visualize size of pixel ----

# 8.1 Make the map --

ggplot() +
  geom_sf(data = grid_reefs, aes(fill = ECOREGION), show.legend = FALSE) +
  geom_sf(data = data_map) +
  coord_sf(x = c(30, 60), y = c(-30, -5))

# 8.2 Save the map --

ggsave(filename = "../figs/09_dataviz-size-pixel.png", height = 6, width = 8)

# 9. Remove useless objects ----

rm(data_reefs, grid_reefs, pixel_total, pixel_monitoring, data_meow)

```

## 7.2 Lollipop

```{r fig.height=15, fig.width=10}

# 1. Transform the data ----

data_descriptors <- data_descriptors %>% 
  mutate(color = "#446cb3",
         name = glue("{ECOREGION} (<b style='color:{color}'>{ECO_CODE_X}</b>)"))

# 2. Make the plots ----

# 2.1 Relative coral reefs surface by ecoregion --

plot_area <- ggplot(data = data_descriptors, aes(x = fct_reorder(name, area_rel), y = area_rel)) +
  geom_point() +
  geom_segment(aes(xend = fct_reorder(name, area_rel), y = 0, yend = area_rel)) +
  coord_flip() +
  theme(axis.text.y = element_markdown()) +
  labs(x = NULL, y ="Area (%)")

# 2.2 Number of surveys by ecoregion --

plot_surveys <- ggplot(data = data_descriptors, aes(x = fct_reorder(name, area_rel), y = n_surveys_rel)) +
  geom_point() +
  geom_segment(aes(xend = fct_reorder(name, area_rel), y = 0, yend = n_surveys_rel)) +
  coord_flip() +
  theme(axis.text.y = element_markdown()) +
  labs(x = NULL, y = "Surveys (%)") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

# 2.3 Number of years with data by ecoregion --

plot_years <- ggplot(data = data_descriptors, aes(x = fct_reorder(name, area_rel), y = n_years_abs)) +
  geom_point() +
  geom_segment(aes(xend = fct_reorder(name, area_rel), y = 0, yend = n_years_abs)) +
  coord_flip() +
  theme(axis.text.y = element_markdown()) +
  labs(x = NULL, y = "Years (n)") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

# 2.4 Number of pixel with data by ecoregion --

plot_pixel <- ggplot(data = data_descriptors, aes(x = fct_reorder(name, area_rel), y = n_percent)) +
  geom_point() +
  geom_segment(aes(xend = fct_reorder(name, area_rel), y = 0, yend = n_percent)) +
  coord_flip() +
  theme(axis.text.y = element_markdown()) +
  labs(x = NULL, y = "Pixel (%)") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

# 3. Assemble the plots ----

plot_grid(plot_area, plot_surveys, plot_years, plot_pixel, nrow = 1, rel_widths = c(1, 0.35, 0.35, 0.35))

# 4. Save the plot ----

ggsave(filename = "../figs/09_ecoregions-data-descriptors.png", height = 15, width = 14)

# 5. Remove useless objects ----

rm(plot_area, plot_surveys, plot_years, plot_pixel)

```

## 7.3 Multivariate approach

### 7.3.1 Analyses

```{r}

# 1. PCA ----

# 1.1 Make the PCA --

data_descriptors_pca <- as_tibble(data_descriptors) %>% 
  select(ECOREGION, n_sites_abs, n_surveys_abs, first_year, n_15_abs, n_years_abs, n_percent)

res_pca <- PCA(data_descriptors_pca[,-1])

# 1.2 Assemble plots --

plot_grid(plot.PCA(res_pca, choix = "ind"), 
          plot.PCA(res_pca, choix = "var"))

# 1.3 Save the plot --

ggsave(filename = "./../figs/10_multivariate-approach-pca.png")

# 2. HCPC ----

# 2.1 Make the HCPC --

res_hcpc <- HCPC(res_pca, nb.clust=5) # Arbitrary selection of the number of cluster 
# (remove nb.clust argument to define the cutting by a click)

# 2.2 Save the plots --

png("./../figs/10_multivariate-approach-hcpc1.png", width = 1600, height = 1600, res = 300)
plot.HCPC(res_hcpc, choice = "tree")
dev.off()

png("./../figs/10_multivariate-approach-hcpc2.png", width = 1600, height = 1600, res = 300)
plot.HCPC(res_hcpc, choice = "map")
dev.off()

```

### 7.3.2 Figures

```{r}

# 1. Graph of variables ----

# 1.1 Extract and transform the data --

data_var <- as_tibble(res_pca$var$coord) %>% 
  bind_cols(var = rownames(res_pca$var$coord), .) %>% 
  select(var, Dim.1, Dim.2)

# 1.2 Make the graph --

ggplot(data = data_var) +
  geom_text_repel(aes(x = Dim.1, y = Dim.2, label = var)) +
  geom_segment(aes(x = 0, y = 0, xend = Dim.1, yend = Dim.2),
                  arrow = arrow(length = unit(0.25, "cm"))) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  lims(x = c(-1, 1), y = c(-1, 1)) +
  geom_circle(data = data.frame(x0 = 0, y0 = 0, r = 1), # From ggforce package
              aes(x0 = x0, y0 = y0, r = r)) +
  coord_fixed() +
  labs(x = paste0("Dim 1 (", round(res_pca$eig[1,2], 2), "%)"), 
       y = paste0("Dim 2 (", round(res_pca$eig[2,2], 2), "%)"))

ggsave(filename = "./../figs/10_multi-approach_graph-var.png", height = 6, width = 6)

# 2. Graph of individuals ----

# 2.1 Extract and transform the data --

data_hcpc_descriptors <- res_hcpc$data.clust %>% 
  mutate(ind = rownames(.))

data_hcpc_cluster <- res_hcpc$call$X %>% 
  mutate(ind = rownames(.))

data_ind <- left_join(data_hcpc_cluster, data_hcpc_descriptors) %>% 
  left_join(data_descriptors, .) %>% 
  st_as_sf() %>% 
  select(ECOREGION, ECO_CODE_X, Dim.1, Dim.2, clust) %>% 
  # Change numbering of clusters to match with logical order for interpretation
  mutate(clust = as.character(clust),
         clust = replace_na(clust, "6"),
         clust = case_when(clust == "1" ~ "5",
                           clust == "2" ~ "4",
                           clust == "4" ~ "2",
                           clust == "5" ~ "1",
                           TRUE ~ clust))

# Palette of colors for each cluster
palette_cluster <- c("1" = "#03a678",
                     "2" = "#68c3a3",
                     "3" = "#19b5fe",
                     "4" = "#eb9532",
                     "5" = "#d64541",
                     "6" = "#6c7a89")

# Get the numbers of ecoregions for each cluster
legend_cluster <- data_ind %>%
  st_drop_geometry() %>% 
  group_by(clust) %>% 
  count()

# Legend labels for each cluster
legend_cluster <- c("1" = glue("<b>1.</b> Very high spatio-temporal coverage (<i>n</i> = 2)"),
                    "2" = glue("<b>2.</b> High spatio-temporal coverage (<i>n</i> = 8)"),
                    "3" = glue("<b>3.</b> Medium spatial, medium temporal coverage (<i>n</i> = 32)"),
                    "4" = glue("<b>4.</b> Medium spatial, low temporal coverage (<i>n</i> = 23)"),
                    "5" = glue("<b>5.</b> Low spatial, low temporal coverage (<i>n</i> = 29)"),
                    "6" = glue("<b>6.</b> No data (<i>n</i> = 12)"))

# 2.2 Make the graph without clusters colors --

ggplot(data = data_ind, aes(x = Dim.1, y = Dim.2, label = ECO_CODE_X)) +
  geom_point(size = 2) +
  geom_text(show.legend = FALSE, size = 3, nudge_x = 0.25) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = paste0("Dim 1 (", round(res_pca$eig[1,2], 2), "%)"), 
       y = paste0("Dim 2 (", round(res_pca$eig[2,2], 2), "%)"))

ggsave(filename = "./../figs/10_multi-approach_graph-ind.png", height = 6, width = 6)

# 2.3 With clusters colors --

ggplot(data = data_ind, aes(x = Dim.1, y = Dim.2, label = ECO_CODE_X, col = clust)) +
  geom_point(show.legend = FALSE, size = 2) +
  geom_text(show.legend = FALSE, size = 3, nudge_x = 0.25) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = paste0("Dim 1 (", round(res_pca$eig[1,2], 2), "%)"), 
       y = paste0("Dim 2 (", round(res_pca$eig[2,2], 2), "%)")) +
  scale_color_manual(values = palette_cluster)

ggsave(filename = "./../figs/10_multi-approach_graph-ind-clust.png", height = 6, width = 6)

# 3. Map with all clusters ----

# 3.1 Make the map --

ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # Clusters
  geom_sf(data = data_ind, aes(fill = clust)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  guides(fill = guide_legend(title.position = "top", title.hjust = 0)) +
  theme_map() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "bold"),
        legend.text = element_markdown()) +
  scale_fill_manual(values = palette_cluster, labels = legend_cluster, name = "Cluster and interpretation")

# 3.2 Save the map --

ggsave(filename = "./../figs/10_multi-approach_map.png", height = 5, width = 12)

# 4. Map by cluster ----

for(i in 1:6){
  
  # 4.1 Make the map --

  ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_meowreefs, fill = "#ecf0f1") +
  # Clusters
  geom_sf(data = data_ind %>% filter(clust == as.character(i)), aes(fill = clust)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  guides(fill = guide_legend(title.position = "top", title.hjust = 0)) +
  theme_map() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "bold"),
        legend.text = element_markdown()) +
  scale_fill_manual(values = palette_cluster, labels = legend_cluster, name = "Cluster and interpretation")

  # 4.2 Save the map --

  ggsave(filename = paste0("./../figs/10_multi-approach_map_cluster-", i,".png"), height = 5, width = 12)

}

# 5. Remove useless objects ----

rm(data_var, data_ind, data_hcpc_descriptors, data_hcpc_cluster, i, legend_cluster)

```

## 7.4 Surveys effort vs reef area

```{r}

# 1. Survey effort vs reef area ----

# 1.1 Make the plot --

ggplot(data = data_descriptors, aes(x = area_rel, y = n_surveys_rel, label = ECO_CODE_X)) +
  geom_point() +
  geom_text_repel() +
  geom_abline(intercept = 0) +
  labs(x = "Relative reef surface (%)", y = "Relative surveys number (%)")

# 1.2 Save the plot --

ggsave(filename = "./../figs/11_survey-effort-vs-reef-area_raw.png", height = 8, width = 8)

# 2. Deviation of survey effort vs reef area ----

# 2.1 Calculate the deviation from expected for surveys --

data_descriptors_effort <- data_descriptors %>% 
  mutate(survey_deviation = n_surveys_rel - area_rel) %>% 
  st_as_sf()

# 2.2 Make the plot --

palette_fill <- c("#d64541", "white", "#03a678")

ggplot(data = data_descriptors_effort, aes(x = area_rel, y = survey_deviation, label = ECO_CODE_X, fill = survey_deviation)) +
  geom_segment(aes(x = area_rel, xend = area_rel, y = 0, yend = survey_deviation), col = "black") +
  geom_point(size = 3, shape = 21, show.legend = FALSE) +
  geom_text_repel(col = "black") +
  geom_hline(yintercept = 0) +
  labs(x = "Relative reef surface (%)", y = "Deviation from expected") +
  lims(y = c(-10, 10)) +
  scale_fill_gradientn(colours = palette_fill, limits = c(-10, 10), breaks = c(-10, -5, 0, 5, 10))

# 2.3 Save the plot --

ggsave(filename = "./../figs/11_survey-effort-vs-reef-area_graph-deviation.png", height = 4, width = 6)

# 3. Map of deviation ----

# 3.1 Make the plot --

ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_descriptors_effort, aes(fill = survey_deviation)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  # Set guide colours
  scale_fill_gradientn(colours = palette_fill, limits = c(-10, 10), breaks = c(-10, -5, 0, 5, 10)) +
  guides(fill = guide_colourbar(title = "Deviation from expected",
                                title.position = "top",
                                barwidth = 18,
                                barheight = 1,
                                frame.colour = "black",
                                frame.linewidth = 1,
                                label.theme = element_text(family = font_choose_map,
                                                           size = 10),
                                ticks.colour = "black")) +  
  theme_map()

# 3.2 Save the map --

ggsave("../figs/11_survey-effort-vs-reef-area_map-deviation.png", height = 6, width = 10)

```

## 7.5 Maps of descriptors

### 7.5.1 Reef area

```{r}

# 1. Create bins ----

hist(data_descriptors$area_rel)

data_descriptors_maps <- data_descriptors %>% 
  mutate(area_rel_cut = cut(area_rel, breaks = c(0, 0.25, 2, 4, 8), include.lowest = TRUE))

# 2. Make the map ----

map_area <- ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_descriptors_maps, aes(fill = area_rel_cut)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  scale_fill_manual(values = palette_fill_2) +
  guides(fill = guide_legend(title.position = "top",
                      title.hjust = 0.5,
                      title = "Relative coral reef area (%)")) +
  labs(title = "AREA") + 
  theme_map2()

# 3. Save the map ----

ggsave(map_area, file = "../figs/10_global-map_reef-area.png", height = 6, width = 10)

```

### 7.5.2 Pixel with data

```{r}

# 1. Create bins ----

hist(data_descriptors$area_rel)

data_descriptors_maps <- data_descriptors %>% 
  mutate(percent_n_cut = cut(n_percent, breaks = c(0, 25, 50, 75, 100), include.lowest = TRUE))

# 2. Make the map ----

map_pixel <- ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_descriptors_maps, aes(fill = percent_n_cut)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  scale_fill_manual(values = palette_fill_2) +
  guides(fill = guide_legend(title.position = "top",
                      title.hjust = 0.5,
                      title = "Percentage of pixel with data")) +
  labs(title = "SPATIAL") + 
  theme_map2()

# 3. Save the map ----

ggsave(map_pixel, file = "../figs/10_global-map_pixel-with-data.png", height = 6, width = 10)

```

### 7.5.3 First year

```{r}

# 1. Create bins ----

hist(data_descriptors$first_year)

data_descriptors_maps <- data_descriptors %>% 
  mutate(first_year_cut = case_when(is.na(first_year) ~ "No data",
                                    first_year >= 1980 & first_year < 2000 ~ "1980 - 1999",
                                    first_year >= 2000 & first_year < 2010 ~ "2000 - 2009",
                                    first_year >= 2010 ~ "> 2010"),
                    first_year_cut = factor(first_year_cut, 
                                            levels = c("1980 - 1999", "2000 - 2009", "> 2010", "No data")))

# 2. Make the map ----

map_first <- ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_descriptors_maps, aes(fill = first_year_cut)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  scale_fill_manual(values = rev(palette_fill_2)) +
  guides(fill = guide_legend(title.position = "top",
                      title.hjust = 0.5,
                      title = "Year with first monitoring")) +
  labs(title = "TEMPORAL") + 
  theme_map2()

# 3. Save the map ----

ggsave(map_first, file = "../figs/10_global-map_first-year.png", height = 6, width = 10)

```

### 7.5.4 Years with data

```{r}

# 1. Create bins ----

hist(data_descriptors$n_years_abs)

data_descriptors_maps <- data_descriptors %>% 
    mutate(n_years_abs_cut = case_when(n_years_abs == 0 ~ "No data",
                                       n_years_abs > 0 & n_years_abs <= 10 ~ "1 - 10",
                                       n_years_abs > 10 & n_years_abs <= 20 ~ "11 - 20",
                                       n_years_abs > 20 ~ "> 21"),
           n_years_abs_cut = factor(n_years_abs_cut, levels = c("No data", "1 - 10", "11 - 20", "> 21")))

# 2. Make the map ----

map_years <- ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_descriptors_maps, aes(fill = n_years_abs_cut)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  scale_fill_manual(values = palette_fill_2) +
  guides(fill = guide_legend(title.position = "top",
                      title.hjust = 0.5,
                      title = "Number of years with data over 1980 - 2020")) +
  labs(title = "TEMPORAL") + 
  theme_map2()

# 3. Save the map ----

ggsave(map_years, file = "../figs/10_global-map_years-with-data.png", height = 6, width = 10)

```

### 7.5.5 Surveys

```{r}

# 1. Create bins ----

hist(data_descriptors$n_surveys_rel)

data_descriptors_maps <- data_descriptors %>% 
    mutate(n_surveys_rel_cut = case_when(n_surveys_rel == 0 ~ "No data",
                                         n_surveys_rel > 0 & n_surveys_rel <= 0.5 ~ "(0,0.5]",
                                         n_surveys_rel > 0.5 & n_surveys_rel <= 1 ~ "(0.5,1]",
                                         n_surveys_rel > 1 ~ "> 1"),
           n_surveys_rel_cut = factor(n_surveys_rel_cut, levels = c("No data", "(0,0.5]", "(0.5,1]", "> 1")))

# 2. Make the map ----

map_surveys <- ggplot() +
  # Background map border
  geom_sf(data = background_map_border, fill = "#56B4E950", color = "grey30", size = 0.5/.pt) +
  # MEOW
  geom_sf(data = data_descriptors_maps, aes(fill = n_surveys_rel_cut)) +
  # Background map
  geom_sf(data = data_map, fill = col_fill_map, col = col_color_map) +
  scale_fill_manual(values = palette_fill_2) +
  guides(fill = guide_legend(title.position = "top",
                      title.hjust = 0.5,
                      title = "Relative number of surveys (%)")) +
  labs(title = "SPATIAL") + 
  theme_map2()

# 3. Save the map ----

ggsave(map_surveys, file = "../figs/10_global-map_relative-surveys.png", height = 6, width = 10)

```

### 7.5.6 Bind plots

```{r fig.height=12, fig.width=10}

# 1. Make the map ----

plot_grid(map_area, NULL, map_pixel, map_surveys, map_first, map_years, nrow = 3, ncol = 2)

# 2. Save the map ----

ggsave(file = "../figs/10_global-map_combined-maps.png", height = 12, width = 10)

```

# 8. Supplementary table

```{r}

data_descriptors %>% 
  select(ECO_CODE_X, ECOREGION, area_abs, area_rel, n_1_abs, n_2_5_abs, n_6_10_abs, 
         n_11_15_abs, n_15_abs, n_surveys_abs, n_years_abs, first_year, n_percent) %>%
  rename(Code = ECO_CODE_X, Name = ECOREGION, Surveys = n_surveys_abs) %>% 
  st_drop_geometry() %>% 
  mutate(area_abs = round(area_abs*1e-6, 1),
         area_rel = round(area_rel, 3),
         n_percent = round(n_percent, 1)) %>%
  write.csv2(., file = "../figs/xx_supplementary-table.csv", row.names = FALSE)

```

# Reproducibility

```{r reprod}

# 1. Reproducibility ----

sessionInfo()

```

---
Jeremy WICQUART | jeremywicquart@gmail.com | `r format(Sys.time())`